\thetitle

### 2008-06-05 paper.tex
More work on Related Work.

More than twenty years of research in analysis and modeling of network traffic have yielded tremendous advances in our understanding of the complex behaviors that emerge from the interaction of millions of humans and computers on the Internet and in local-area networks (\caps{LAN}s).
Both traffic analysis and realistic workload generation, however, remain active areas of research with many unanswered questions.
One of the major challenges of modeling full-network traffic is the multilevel interdependence of packet, flow\flownote and node behaviors.
Three types of simplifying assumptions are commonly applied in both analysis and generation:
\begin{enumerate}
\item \textbf{Uniform:} that traffic properties are uniformly distributed. For example, assuming that  all nodes in a network have equal probability of being the source or destination of each flows.
Another common example is the assumption that all packet sizes and inter-packet intervals are the same; this is commonly known as the ``constant bit-rate'' (\caps{CBR}) model.
\item \textbf{Marginal:} that traffic properties share a single, common \emph{marginal}\marginalnote distribution across the entire network, and individual values are drawn independently and unconditionally from this shared distribution.
For example, assuming that a single distribution of packet sizes describes all flows in a network, as compared to equipping each flow in the network with its own specific distribution.
\item \textbf{Conditional:} that traffic behaves marginally, but only within equivalence classes defined by certain conditions.
The assumption, for example, that all flows emanating from a common source node share the same packet size distribution is a conditional model, conditioned on source nodes.
\end{enumerate}
We would be arguing against a straw man if we implied that researchers who apply these assumptions actually believe them to be true.
They clearly are not: flows have drastically different packet size distributions; different nodes initiate and receive drastically different numbers of packets and flows.
Simply considering intuitive examples of common behavior illustrates this amply:
downloading a large file \latin{vs.} typing in an \caps{SSH} session;
BitTorrent users \latin{vs.} occasional email checkers.

@@

\subsection{Traffic Generation}

Paxson and Floyd observed~\cite{Paxson95} that the interaction between endpoint behavior and the network conditions is inherently \textit{closed-loop} in the sense that the exhibited behavior is governed by a closed feedback loop.
Traffic models typically attempt to preserve the closed-loop behavior of network traffic.
In the case of most non-\caps{TCP} applications, this feedback does not occur: an open-loop model suffices.
Hern\'andez-Campos's dissertation~\cite{Hernandez06:dissertation} addresses in great depth how to model the feedback between the network and applications for most \caps{TCP} applications.
Hern\'andez-Campos shows how to abstract an application's behavior from a \caps{TCP} trace and replay it in a manner that accurately reflects how most common applications would react in response to different network conditions.
This research has subsequently been turned into a \caps{TCP} flow generation tool called \tmix~\cite{Weigle06}.
It is important to note that {\tmix} does not provide models for application behaviors or for mixtures of applications in networks.
Rather, it provides the ability to reproduce accurate closed-loop dynamics for a single \caps{TCP} flow \emph{given} an application behavior model as input.
Our work complements this perfectly: this paper provides the means to analyze traffic and generate realistic whole-network application workloads which can then serve as the necessary input to \tmix.

@@

% There is a large and diverse body of work on traffic analysis, modeling, and generation~\cite{Paxson95,Paxson96,Sommers04,Avallone04,Hernandez06:dissertation}. We are only able to discuss a small, but hopefully representative sampling of this work. Almost all of the traffic generation work has focused on wide-area Internet backbone traffic.
%Because of the comparatively unlimited capacity of wired networks, traffic workload generally does not offer interesting technical challenges at the edge of the network.
The two most prominent general traffic generation frameworks are Harpoon and \caps{D-ITG}. Harpoon~\cite{Sommers04} uses a traffic trace for self-training, and can subsequently generate synthetic traffic with certain statistical properties based on the original trace.
The properties reproduced are the empirical distributions of the following: ``file size, inter-connection time, source and destination IP ranges, number of active sessions.''
Harpoon exemplifies the uniform and marginal modeling approaches.
Source and destination nodes are treated marginally: an empirically derived marginal frequencies of sources and destination nodes (independently) are used to choose endpoints for each flow.
File sizes (\latin{i.e.} total flow sizes) are likewise treated marginally: each flow's total byte  count is sampled from a marginal distribution of flow sizes.
\caps{UDP} flows are modeled as being constant bit-rate (\latin{i.e.} having uniform packet behavior), and \caps{TCP} flows are all treated as file transfers, making them effectively uniform with bandwidth limited by simulated \caps{TCP} dynamics.
Inter-connection times (\latin{i.e.} per node-pair inter-flow intervals) are sampled from a single empirically derived marginal distribution.
While Harpoon may be sufficiently realistic for the intended purpose of generating Internet backbone traffic, it is difficult to know this:
generated traffic is only compared with trace traffic using metrics that are explicitly specified in the models.
Realism is never double-checked using derived statistical metrics or actual performance metrics.
For the purpose of generating local-area traffic for wireless experiments, it is certainly not sufficiently realistic. Previous research~\cite{Karpinski07:cbr-failure} has shown that \caps{CBR} packet behavior models, like those used in Harpoon, drastically distort vital performance metrics at all levels of the protocol stack in wireless network simulations. This paper gives equal cause for concern about marginal traffic models.

% There is no criterion proposed to determine whether these properties characterize the original traffic adequately---we can only hope that this approximation is good enough. For many purposes, it likely is sufficient; in particular, it is probably appropriate for the intended use in generating traffic for Internet backbone simulations. Wireless networks, however, are particularly sensitive to workload conditions, and sampling from a limited set of empirical distributions does not suffice to reproduce realistic network-wide traffic.

\caps{D-ITG}~\cite{Avallone04} generates flows using an independent sampling model for packet sizes and inter-packet intervals. The framework contains pre-made models for several common types of Internet traffic. The focus of this project, however, is on providing the infrastructure to generate very large volumes of synthetic Internet-like backbone traffic. No analysis is provided for determining the realism of traffic mixes, or for choosing flow endpoints realistically. In wireless networks, these factors are of crucial importance to performance, and cannot be overlooked. Both Harpoon and \caps{D-ITG} provide excellent traffic generation platforms, but do not provide a systematic framework for understanding or reproducing realistic whole-network workload in the wireless setting.

>>>>

More than twenty years of research in analysis and modeling of network traffic have yielded tremendous advances in our understanding of the complex behaviors that emerge from the interaction of millions of humans and computers on the Internet and in local-area networks (\caps{LAN}s).
Both traffic analysis and realistic workload generation, however, remain active areas of research with many unanswered questions.
One of the major challenges of modeling whole-network traffic is the interdependence of packet, flow\flownote and node behaviors.
Three types of simplifying assumptions are commonly applied in both analysis and generation:
\begin{enumerate}
\item \textbf{Uniform:} that traffic properties are uniformly distributed. For example, assuming that  all nodes in a network have equal probability of being the source or destination of each flow.
Another common example is the assumption that all packet sizes and inter-packet intervals are the same; this is commonly known as the ``constant bit-rate'' (\caps{CBR}) model.
\item \textbf{Marginal:} that traffic properties share a single, common \emph{marginal}\marginalnote distribution across the entire network, and individual values are drawn independently and unconditionally from this shared distribution.
For example, assuming that a single distribution of packet sizes describes all flows in a network, as compared to equipping each flow in the network with its own specific distribution.
\item \textbf{Conditional:} that traffic behaves marginally, but only within equivalence classes defined by certain conditions.
The assumption, for example, that all flows emanating from a common source node share the same packet size distribution is a conditional model, conditioned on source nodes.
\end{enumerate}
We would be arguing against a straw man if we implied that researchers who apply these assumptions actually believe them to be true.
They clearly are not: flows have drastically different packet size distributions; different nodes initiate and receive drastically different numbers of packets and flows.
Simply considering intuitive examples of common behavior illustrates this amply:
downloading a large file \latin{vs.} typing in an \caps{SSH} session;
BitTorrent users \latin{vs.} occasional email checkers.

@@

While this work is not inherently specific to wireless networks, since the modeling techniques developed here equally applicable to traffic in wired networks, the results are of fundamental importance to wireless research. Realistic traffic models have a drastic impact on the experimental results in wireless networks~\cite{Karpinski07:realism,Karpinski07:cbr-failure}. This is primarily because wireless networks are locally resource constrained. Compare this with the typical situation in wired \caps{LAN}s, where gigabit Ethernet and modern routers and switches can easily handle all but the most intense levels of traffic. If a wired network operator expects larger volumes of traffic on a regular basis, the solution is to upgrade the hardware, not to change the network protocols. In wireless networks, on the other hand, the physical medium cannot be upgraded: protocols must be improved to make better usage of the medium. Being able to perform experiments with realistic workload is a vital part of the ongoing process of improving wireless protocols. Without realistic workload models for experiments, performance comparisons may be misleading or even completely wrong~\cite{Karpinski07:cbr-failure}.% Better traffic models are necessary to avoid stumbling in the dark while attempting to improve wireless technologies.

Paxson and Floyd observed~\cite{Paxson95} that the interaction between endpoint behavior and the network conditions is inherently \textit{closed-loop} in the sense that the exhibited behavior is governed by a closed feedback loop.
% Traffic models typically attempt to preserve the closed-loop behavior of network traffic.
In the case of most non-\caps{TCP} applications, this feedback does not occur: an open-loop model suffices.
Hern\'andez-Campos' dissertation~\cite{Hernandez06:dissertation} addresses in great depth how to model the feedback between the network and applications for typical \caps{TCP} applications.
Hern\'andez-Campos shows how to abstract an application's behavior from a \caps{TCP} trace and replay it in a manner that accurately reflects how typical applications react in response to different network conditions.
This research has subsequently been turned into a \caps{TCP} flow generation tool called \tmix~\cite{Weigle06}.
The acronym is somewhat misleading: {\tmix} does not provide models for application behaviors or for mixtures of applications in networks.
Rather, it provides the ability to reproduce accurate closed-loop dynamics for a single \caps{TCP} flow \emph{given} an application behavior model as input.
Our research complements this perfectly: we provide the means to analyze traffic and generate realistic whole-network application mixtures which can then serve as the necessary input to \tmix.

@@

% There is a large and diverse body of work on traffic analysis, modeling, and generation~\cite{Paxson95,Paxson96,Sommers04,Avallone04,Hernandez06:dissertation}. We are only able to discuss a small, but hopefully representative sampling of this work. Almost all of the traffic generation work has focused on wide-area Internet backbone traffic.
%Because of the comparatively unlimited capacity of wired networks, traffic workload generally does not offer interesting technical challenges at the edge of the network.

\newfootnote{\harpoonnote}{Even this is unclear: generated traffic is only compared with trace traffic using the very same metrics that are explicitly specified as part of the model. Unsurprisingly, the distributions of sampled metrics closely match the distributions those metrics were sampled from. Realism is not double-checked using other statistical metrics or actual performance metrics.}

The two most prominent general traffic generation frameworks are Harpoon and \caps{D-ITG}. Harpoon~\cite{Sommers04} uses a traffic trace for self-training, and can subsequently generate synthetic traffic with certain statistical properties based on the original trace.
The properties reproduced are the empirical distributions of the following: ``file size, inter-connection time, source and destination IP ranges, number of active sessions.''
Harpoon exemplifies the uniform and marginal modeling approaches.
Source and destination nodes are treated marginally: empirically derived marginal frequencies of sources and destination nodes (independently) are used to choose endpoints for each flow.
File sizes (\latin{i.e.} total flow sizes) are likewise treated marginally: each flow's total byte  count is sampled from a marginal distribution of flow sizes.
\caps{UDP} flows are modeled as being constant bit-rate (\latin{i.e.} having uniform packet behavior), and \caps{TCP} flows are all treated as file transfers, making them effectively uniformly behaved with bit-rate determined by \caps{TCP} dynamics.
Inter-connection times (\latin{i.e.} per node-pair inter-flow intervals) are sampled from a single empirically derived marginal distribution.
While Harpoon may be sufficiently realistic for the intended purpose of generating Internet backbone traffic,\harpoonnote
for the purpose of generating local-area traffic for wireless experiments, it is not sufficiently realistic.
Previous research~\cite{Karpinski07:cbr-failure} has shown that \caps{CBR} packet behavior models used in Harpoon drastically distort vital performance metrics at all levels of the protocol stack in wireless network simulations.
This paper shows that marginal traffic models also severely misrepresent performance metrics.

Like Harpoon, \caps{D-ITG}~\cite{Avallone04} exemplifies the common uniform and marginal modeling assumptions.
Packet behavior is modeled using a handful of standard statistical distributions (``constant, uniform, normal, cauchy, pareto, exponential, etc.'') for packet size and inter-packet interval distributions, from which packets are generated using repeated independent sampling.
Guidance is not provided, however, for choosing among these distributions, choosing parameter values for individual flows, or selecting a mixture of distributions for flows across an entire network.
Moreover, the mechanism for choosing end-points and start and end times for flows is not described; presumably it is explicitly specified by the user. 
The focus of this project is primarily on the engineering challenge of generating very large volumes of synthetic Internet-like backbone traffic and injecting this traffic into real networks in a distributed fashion.
While \caps{D-ITG} provides the mechanism for generating large volumes of network traffic, it does not provide any real guidance as to what traffic to generate.

% The focus of this project, however, is on providing the infrastructure to generate very large volumes of synthetic Internet-like backbone traffic. No analysis is provided for determining the realism of traffic mixes, or for choosing flow endpoints realistically. In wireless networks, these factors are of crucial importance to performance, and cannot be overlooked. Both Harpoon and \caps{D-ITG} provide excellent traffic generation platforms, but do not provide a systematic framework for understanding or reproducing realistic whole-network workload in the wireless setting.
